Stack Safety is a Security Property

### 2021-04-17 tex/main.tex
Pass up to 7

\paragraph*{What is a ``call''?}  One significant challenge in formulating our
properties arises from the fact that we want to define stack safety at the
machine-code level for a variety of ISAs, including RISCy \sna{RISCy?} ones.  In
particular, there may be no single ``call'' instruction that does all the
required manipulation of the \(\PCname\) and stack pointer. Instead, there
might be just a simple branch instruction that stores the old \(\PCname\) in
a register, with the rest of the call sequence performed by other
instructions either before or after the jump.  Similarly, there may be no
``return'' macro-instruction, but rather a ``return sequence'' involving an
indirect branch at some point.  Thus, the same instruction opcode may be
used in some places as part of a call or return sequence and in other places
for other purposes.

@@

This form of confidentiality we will term {\em stepwise confidentiality}, defined informally
as: for any state at the entry to a function, for any variant over the values in the stack,
the primary state and its variant will step in lockstep until they return. To step in lockstep
means that if both states step---that is, neither the primary nor the variant failstops---any
{\em system component} (such as memory cell, register, program counter) that changes
in one step has a matching value at the end of the other step. When we later
extend the model with a notion of observable output, values that are emitted will need
to match as well.

@@

In particular, we wish to formulate properties around consecutive subtraces
in which the context obeys a given predicate. This will commonly be used to extract
individual calls from a trace by taking subtraces in which the stack is
at or below a given depth. If \(f\) is still a predicate on states, then
we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
``\(\mpcstatename\) segmented by \(f\) yields a sub-trace \(\MPCT\).''
\bcp{Throughout, we could save space and improve beauty by removing some of the
vertical space between inference rules.}

@@

The rules for \(\stepsto_C\) are given inductively, and apply in the listed order.
A call annotation tells us to push the return target onto the return stack and
seal all components with the depth of the previous return stack. If we aren't making a call,
we identify a return by looking ahead and checking if we're about to step to a state
that matches the topmost return target. If neither is true, the context remains unchanged.
Recall that a ``code map'' \(\codemap\) is a predicate on addresses.
\rb{Technically a partial function from addresses to annotations,
  though its domain can function as a predicate.}

@@

In other words, stepwise stack confidentiality states that if, at the start of the call,
we vary all stack locations that are supposed to be secret from the callee
(which in our simple setting is the entirety of the stack),
the resulting trace remains in lockstep with the original trace
until either both return together or one failstops. This implies that
if the callee reads a secret location, it must failstop.\apt{right?}
%
Note that \(\MPCT\) may contain nested calls; these must also respect the secrecy
of the components in \(\components\), and in addition their subtraces are also in
\(\MPCT^*\), and therefore confidentiality must hold on them separately. \(\NPCT\)
will contain these calls and returns as well, so its return stack pushes and pops
in lockstep until it breaks at the final return.

@@

In this section we explore how an existing enforcement mechanism, {\em
  Depth Isolation}~\citep{DBLP:conf/sp/RoesslerD18}, implements
the formal stack-safety property described in the previous
section. (In fact, it enforces a slightly stronger, inductive version
of that property which checks for integrity and confidentiality violations at
every step of the callee's execution trace.) We validate this claim
with a property-based testing framework that can quickly uncover
inconsistencies between the enforcement mechanism and the stack-safety
property.

\subsection{A Conservative Stack-Safety Enforcement Mechanism}
%
The enforcement policy of \citeauthor{DBLP:conf/sp/RoesslerD18} relies
on a programmable, tag-based reference monitor that runs alongside the
program,
allowing fine-grained manipulation of metadata tags to encode so-called
\emph{micro-policies}~\citep{pump_oakland2015}.
Every value in
memory and registers (including the $\PCname$)
is enriched with an abstract metadata tag, which can represent
arbitrary information about the value. A micro-policy is then defined as a
set of tags and a
collection of software-defined rules.
At each step of the machine, the relevant rule is applied to the tags on
the instruction's inputs ($\PCname$, registers, memory) and on the
instruction opcode itself, producing
one of two outcomes: either the instruction is permitted to execute (and
generates tags for the result of the operation and the new $\PCname$), or
the machine fail-stops with a policy violation.
\citeauthor{pump_oakland2015}\bcp{additional citations: the noninterference
  paper, Nick and Andre's papers, ...} show that a wide range of
micro-policies can be defined using this scheme.

@@

The micro-policy maintains tags of the form $\tagStackDepth{n}$ for
each stack location belonging to the stack frame at activation depth
$n$, or the tag $\tagNoDepth$ for stack locations that are currently
unused (but could potentially be used in the future).
%
The $\PCname$ is tagged similarly as $\tagPCDepth{n}$.
%
Initially, the entire stack is considered unused (and therefore tagged
$\tagNoDepth$), and the {\PCname} has tag $\tagPCDepth{0}$.
%
During normal execution, the micro-policy rules only permit load and
store operations when the target memory is tagged {\em with the same
  depth} as the current {\PCname} tag.

@@

We validate the micro-policy of the previous section by
systematically testing that it satisfies eager stack integrity and
confidentiality. We use a Coq specification of the RISC-V
architecture
\footnote{Available at \url{https://github.com/mit-plv/riscv-coq}.}
\leo{translated from Haskell via hs-2-coq?\apt{not worth saying since
  the MIT people did it}}  and extend it with a
runtime monitor implementing the stack-safety
micro-policy.\footnote{This code will be made public. \leo{More?}} We
  choose the Coq proof assistant as the setting for our implementation
  for two reasons: first, to ensure that our coinductive trace
  definitions are well-formed, and reason about them\leo{Do we prove
    anything interesting anymore?}, and to leverage the power of the
  QuickChick property-based testing framework~\citep{Pierce:SF4}.

@@

We extend this technique to keep track of the control flow behavior of
the program being generated: each time a call or return sequence is
generated, we ensure that the appropriate policy tags and code
annotations are set for the entry or return points. At the same time,
we sometimes allow the generation to relax those constraints,
introducing potentially ill-formed flows: this causes our programs
to failstop when executed in conjunction with the policy monitor,
but also allows for revealing errors in our setup.

@@

Naturally, testing also revealed a number of errors in our
implementation of the enforcement mechanism (the original was written in C++
and targeted ARM machine code;
%\bcp{right?}\leo{yeah}
we re-implemented it in Coq targeting RISC-V).  These errors range
from trivial typos to ones that require an intriguingly complex setup
to appear.  The most interesting bug (included in the table as row
{\em HEADER\_NO\_INIT}) was that, on our first try, the blessed call
sequence %/policy combination\apt{??}
did not initialize all locations for the
newly allocated stack frame correctly, but left some of them as
$\tagNoDepth$. This allowed for a potential integrity violation, but
only if a rather complicated sequence of events occured.
The smallest counterexample requires calling a function {\tt f},
which writes into an uninitialized location $l$ (which is always fine\apt{is it??? in the strict policy?}, and then
calls a further function {\tt g} (which should have the effect of sealing $l$
for integrity purposes). Then {\tt g} attempts to write to $l$;
this is allowed because the enforcement mechanism still has
$l$ tagged as $\tagNoDepth$, but violates the integrity property on {\tt f}'s data.
\sna{I believe what went wrong was that we were off-by-one in {\tt main}'s initialization,
  and the write from {\tt f} was already a violation.}

>>>>

\paragraph*{What is a ``call''?}  One significant challenge in formulating our
properties arises from the fact that we want to define stack safety at the
machine-code level for a variety of ISAs, including RISCy \sna{RISCy?}\leo{RISC-like maybe?} ones.  In
particular, there may be no single ``call'' instruction that does all the
required manipulation of the \(\PCname\) and stack pointer. Instead, there
might be just a simple branch instruction that stores the old \(\PCname\) in
a register, with the rest of the call sequence performed by other
instructions either before or after the jump.  Similarly, there may be no
``return'' macro-instruction, but rather a ``return sequence'' involving an
indirect branch at some point.  Thus, the same instruction opcode may be
used in some places as part of a call or return sequence and in other places
for other purposes.

@@

We will term this form of confidentiality as {\em stepwise confidentiality}. Informally,
for any state at the entry to a function, for any variant over the values in the stack,
the primary state and its variant will step in lockstep until they return. To step in lockstep
means that if both states step---that is, neither the primary nor the variant failstops---any
{\em system component} (such as memory cell, register, program counter) that changes
in one step has a matching value at the end of the other step. When we later
extend the model with a notion of observable output, values that are emitted will need
to match as well.

@@

In particular, we wish to formulate properties around consecutive subtraces
in which the context obeys a given predicate. This will commonly be used to extract
individual calls from a trace by taking subtraces in which the stack is
at or below a given depth. If \(f\) is still a predicate on states, then
we define \(\mpcstatename \hookrightarrow \MPCT \langle f \rangle\), pronounced
``\(\mpcstatename\) segmented by \(f\) yields a sub-trace \(\MPCT\).''
\ifspace{\bcp{Throughout, we could save space and improve beauty by removing some of the
vertical space between inference rules.}\fi

@@

The rules for \(\stepsto_C\) are given inductively, and apply in the listed order.
A call annotation tells us to push the return target onto the return stack and
seal all components with the depth of the previous return stack. If we aren't making a call,
we identify a return by looking ahead and checking if we're about to step to a state
that matches the topmost return target. If neither is true, the context remains unchanged.
Recall that a ``code map'' \(\codemap\) is partial function, and we can use its domain
as a predicate on addresses.

@@

In other words, stepwise stack confidentiality states that if, at the start of the call,
we vary all stack locations that are supposed to be secret from the callee
(which in our simple setting is the entirety of the stack),
the resulting trace remains in lockstep with the original trace
until either both return together or one failstops. This implies that
if the callee reads a secret location, it must failstop.%\apt{right?}\leo{yes}
%
Note that \(\MPCT\) may contain nested calls; these must also respect the secrecy
of the components in \(\components\), and in addition their subtraces are also in
\(\MPCT^*\), and therefore confidentiality must hold on them separately. \(\NPCT\)
will contain these calls and returns as well, so its return stack pushes and pops
in lockstep until it breaks at the final return.

@@

In this section we explore how an existing enforcement mechanism, {\em
  Depth Isolation}~\citep{DBLP:conf/sp/RoesslerD18}, implements
the formal stack-safety property described in the previous
section. In fact, it enforces a slightly stronger, inductive version
of that property which checks for integrity and confidentiality violations at
every step of the callee's execution trace. We validate this claim
with a property-based testing framework that can quickly uncover
inconsistencies between the enforcement mechanism and the stack-safety
property.

\subsection{A Conservative Stack-Safety Enforcement Mechanism}
%
The enforcement policy of \citeauthor{DBLP:conf/sp/RoesslerD18} relies
on a programmable, tag-based reference monitor that runs alongside the
program,
allowing fine-grained manipulation of metadata tags to encode so-called
\emph{micro-policies}~\citep{pump_oakland2015}.
Every value in
memory and registers (including the $\PCname$)
is enriched with an abstract metadata tag, which can represent
arbitrary information about the value. A micro-policy is then defined as a
set of tags and a
collection of software-defined rules.
At each step of the machine, the relevant rule is applied to the tags on
the instruction's inputs ($\PCname$, registers, memory) and on the
instruction opcode itself, producing
one of two outcomes: either the instruction is permitted to execute (and
generates tags for the result of the operation and the new $\PCname$), or
the machine fail-stops with a policy violation.
Existing work \citep{TestingNI:ICFP,pump_oakland2015, DBLP:conf/sp/RoesslerD18}
%\bcp{additional citations: the noninterference paper, Nick and Andre's papers, ...}
has shown that a wide range of
micro-policies can be defined using this scheme.

@@

The micro-policy maintains tags of the form $\tagStackDepth{n}$ for
each stack location belonging to the stack frame at activation depth
$n$, or the tag $\tagNoDepth$ for stack locations that are currently
unused (but could potentially be used in the future).
%
The $\PCname$ is tagged similarly as $\tagPCDepth{n}$.
%
Initially, the entire stack is considered unused (and therefore tagged
$\tagNoDepth$), and the {\PCname} has tag $\tagPCDepth{0}$.
%
During normal execution, the micro-policy rules only permit load and
store operations when the target memory is tagged {\em with the same
  depth} as the current {\PCname} tag, and store operations also when the
target memory is $\tagNoDepth$.

@@

We validate the micro-policy of the previous section by
systematically testing that it satisfies eager stack integrity and
confidentiality. We use a Coq specification of the RISC-V
architecture\footnote{Available at \url{https://github.com/mit-plv/riscv-coq}.}
%\leo{translated from Haskell via hs-2-coq?\apt{not worth saying since the MIT people did it}}
and extend it with a
runtime monitor implementing the stack-safety
micro-policy.\footnote{This code is {\em redacted for review}.} We
  chose the Coq proof assistant as the setting for our implementation
  for two reasons: first, to ensure that our coinductive trace
  definitions are well-formed, and reason about them\leo{Do we prove
    anything interesting anymore?}, and to leverage the power of the
  QuickChick property-based testing framework~\citep{Pierce:SF4}.

@@

We extend this technique to keep track of the control flow behavior of
the program being generated: each time a call or return sequence is
generated, we ensure that the appropriate policy tags and code
annotations are set for the entry or return points. At the same time,
we allow the generation to sometimes relax those constraints,
introducing potentially ill-formed flows: this causes our programs
to failstop when executed in conjunction with the policy monitor,
but also allows for revealing errors in our setup.

@@

Naturally, testing also revealed a number of errors in our
implementation of the enforcement mechanism (the original was written in C++
and targeted ARM machine code;
%\bcp{right?}\leo{yeah}
we re-implemented it in Coq targeting RISC-V).  These errors range
from trivial typos to ones that require an intriguingly complex setup
to appear.  The most interesting bug (included in the table as row
{\em HEADER\_NO\_INIT}) was that, on our first try, the blessed call
sequence %/policy combination\apt{??}
did not initialize all locations for the
newly allocated stack frame correctly, but left some of them as
$\tagNoDepth$. This allowed for a potential integrity violation, but
only if a rather complicated sequence of events occured.
The smallest counterexample requires calling a function {\tt f},
which writes into an uninitialized location $l$,
% (which is always fine\apt{is it??? in the strict policy?}
and then
calls a further function {\tt g} (which should have the effect of sealing $l$
for integrity purposes). Then {\tt g} attempts to write to $l$;
this is allowed because the enforcement mechanism still has
$l$ tagged as $\tagNoDepth$, but violates the integrity property on {\tt f}'s data.
%\sna{I believe what went wrong was that we were off-by-one in {\tt main}'s initialization,
%  and the write from {\tt f} was already a violation.}
